\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{parskip}

\begin{document}

I'm interested in engineering roles at HRT, particularly the Research Engineer position. My background is a mix of optimization research and building production systems, and I'm open to wherever that's most useful.

During my PhD at Georgia Tech, I built safety verification tooling for the Robotarium, a multi-robot testbed that's now supported 16,000+ experiments. My research was in optimization and control theory: mixed-integer and convex programming for trajectory planning, control barrier functions for safety guarantees. The math feels relevant to trading: constrained optimization, real-time decisions under uncertainty, balancing competing objectives.

Since then I've built two startups. At Pytheia, I built pricing optimization and demand prediction systems that we sold to enterprise customers, bootstrapping to \$300k ARR. At Roostr, as sole production engineer, I built an end-to-end agentic system: hybrid LLM + deterministic pipelines, Python/FastAPI backend, and custom tooling for rapid debugging. The system contributed to approximately \$1M in incremental monthly revenue for one customer.

I read your blog post on engineering and interviewing at HRT. The emphasis on finding people who are ``not only brilliant, but also open, curious, kind, practical, and actually enjoy working'' resonates. I genuinely love this work---building systems, solving hard problems, making things better. That's not something I have to manufacture.

\vspace{1em}
Mark Mote

\end{document}
